Stack Safety is a Security Property

### 2021-04-16 tex/main.tex

% BCP: This will intentionally fail when someone adds the real intended
% definition of \WF!
\newcommand{\WF}{\mathit{WF}}

@@

In this section we examine how an existing enforcement mechanism,
Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, can implement
the formal stack-safety property described in the previous section: it in
fact enforces a stronger, inductive version of that property which strengthens the
local state encapsulation component by checking for integrity and
confidentiality violations at every step of the callee's execution trace. \bcp{Examples?}

@@

\item From callee back to caller, when execution is about to reach the return point
  $\ret{c}$
  corresponding to the call at $c$.  The exit sequence retags all locations in the function
  frame with $\tagNoDepth$. At the actual return instruction, the
  $\PCname$ tag is decremented.
  \rb{Problem: this is currently the first mention of $\ret{\cdot}$}

@@

Note that in this policy both the entry and exit sequences
must write to each element of
the frame, which can be quite expensive, especially for programs that
allocate large but sparsely populated frames; we return to this point
in \cref{sec:lazy}.
%
For example, the program in \cref{fig:passingsimple} manages the stack
in this way. Here, the call pseudo-instruction to {\tt f} in line 5
hides a blessed sequence of instructions that allocate space to save
its return address to {\tt main} and seal the protected part of the
stack before {\tt f}'s body starts executing at 10. Conversely, the
return instruction in line 15 restores and jumps to the return address
and deallocates its frame space, and unseals the corresponding parts
of the stack. The function itself allocates and deallocates its own
private frame space in instructions 10 and 14. Additionally, unlike in
the simplified presentation of the policy, the caller shares the
arguments passed to the callee through the stack, and for this creates
and initializes (3--4) and destructs (6) the argument part of the
callee's stack frame.
%
%\rb{Note that we have also considered and implemented more flexible schemes
%  based on allowing operations above the current activation depth. How do we put
%  these together?}
%\rb{Do we want to discuss variations involving frame pointers or more implicit
%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
%\rb{Where to talk about the role of the compiler?}

@@

>>>>

@@

\leo{I am not sure we've hit the correct level of detail in this section.}

In this section we detail an existing enforcement mechanism,
Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements
the formal stack-safety property described in the previous section; in
fact, it enforces a stronger, inductive version of that property which strengthens the
local state encapsulation component by checking for integrity and
confidentiality violations at every step of the callee's execution trace. \bcp{Examples?}

@@

\item From callee back to caller, when execution is about to reach a
  return point.  The exit sequence retags all locations in the
  function frame with $\tagNoDepth$. At the actual return instruction,
  the $\PCname$ tag is decremented. \leo{Not sure the actual policy
    does that now (retag unused). But I don't think it's a
    confidentiality or integrity violation.}

@@

\leo{I'm not a fan of the next paragraph.}
Note that in this policy both the entry and exit sequences
must write to each element of
the frame, which can be quite expensive, especially for programs that
allocate large but sparsely populated frames; we return to this point
in \cref{sec:lazy}.
%
For example, the program in \cref{fig:passingsimple} manages the stack
in this way. Here, the call pseudo-instruction to {\tt f} in line 5
hides a blessed sequence of instructions that allocate space to save
its return address to {\tt main} and seal the protected part of the
stack before {\tt f}'s body starts executing at 10. Conversely, the
return instruction in line 15 restores and jumps to the return address
and deallocates its frame space, and unseals the corresponding parts
of the stack. The function itself allocates and deallocates its own
private frame space in instructions 10 and 14. Additionally, unlike in
the simplified presentation of the policy, the caller shares the
arguments passed to the callee through the stack, and for this creates
and initializes (3--4) and destructs (6) the argument part of the
callee's stack frame.
%
%\rb{Note that we have also considered and implemented more flexible schemes
%  based on allowing operations above the current activation depth. How do we put
%  these together?}
%\rb{Do we want to discuss variations involving frame pointers or more implicit
%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
%\rb{Where to talk about the role of the compiler?}

@@

We validate the micro-policy of the previous section by systematically
testing that it satisfies eager stack integrity and
confidentiality. We used a Coq specification of the RISC-V
architecture \leo{cite: https://github.com/mit-plv/riscv-coq} and
extended it with a runtime monitor implementing the stack-safety
micro-policy. \leo{Footnote: code will be made public} We chose the
Coq proof assistant as the setting for our implementation for two
reasons: first, to ensure that our coinductive trace definitions are
well-formed, and reason about them\leo{Do we prove anything
  interesting anymore?}, and to leverage the power of the QuickChick
property-based testing framework.\leo{cite}

To use QuickChick, we built generators that produce:
\begin{enumerate}
\item An initial RISC-V machine state, including most notably
  the program to be executed.
\item An initial policy state, tagging instructions corresponding to
  ``blessed'' call or return sequences appropriately, while marking
  all potential stack locations as $\tagNoDepth$.
\item A code map and a stack map, accurately characterizing the behavior
  of callers and callees.
\end{enumerate}
%
That is, our generators need to produce not only a low-level program
that exhibits interesting control-flow patterns, but also policy
meta-information and code annotations that would normally be produced
by the compiler.

\paragraph*{Generator Overview}

To write such generators we build on the work of \leo{ICFP 2013},
where they introduced {\em generation by execution} to produce
interesting programs. Generation by execution receives as an input a
partially instantiated machine state, and attempts to generate an
instruction (or a sequence of instructions such as a ``blessed''
sequence) that makes sense locally (e.g., jumps to a potentially valid
code location or loads from a potentially valid stack location). Then,
we step the machine and repeat the process until we generate or
execute some target number of instructions (or if we reached a point
where the machine can't step anymore).

We extend this technique to keep track of the control flow behavior of
the program being generated: each time a call or return sequence is
generated, we ensure that the appropriate policy tags and code
annotations are set for the entry or return points. At the same time,
we sometimes allow the generation to relax those constraints,
introducing potentially ill-formed flows: that causes our programs
to failstop when executed in conjunction with the policy monitor,
but also allows for revealing errors in our setup.

\paragraph*{Findings}

To evaluate the effectiveness of our testing, we turn to {\em mutation
  testing}\leo{cite}. In mutation testing, we inject errors
(mutations) in a program that should cause the property of interest
(here, stack safety) to fail, and ensure that the testing framework
can find them. The bugs we used for our evaluation were both
artificially generated (by deliberately weakening the micro-policy in
ways that should break enforcement), or actual bugs that we discovered
through testing in our implementation. We elaborate on some such bugs
below.

The mean-time-to-failure (MTTF) for various bugs can be found in
\leo{Table with numbers?}. All experiments were run in a \leo{Machine
  spec}. \leo{Something like: There is really no baseline to compare
  this against, but include it here for the future/reasons?}

The artificially injected bugs focus on weakening the tag propagation
and checking mechanism. \leo{Mention them all?}
%
For example, when loading from a stack location, the correct
micro-policy needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the one of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth. We can relax
that restriction by not checking the depth equality, or not checking
the tags at all, causing bugs of different difficulty.
%
Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing these constraints also causes violations to the
property.

Using our testing framework, we also found various errors in our
implementation, ranging from trivial typos to ones that required an
intriguingly complex setup to appear. Most notably, the initial
``blessed'' call sequence/policy combination did not initialize all
locations for the newly allocated stack frame correctly, but left some
of them as $\tagNoDepth$. This allows for a potential integrity
violation, but only if a rather complicated sequence of events
occured: \leo{call - write - call (seal) - return - call - write?}
